---
title: 'Lin et al.: AAV Microglia'
subtitle: 'Correlation between Transgene and Marker Gene (Supplementary Figure 3)'
author: "Ruiyu Ray Wang"
date: "`r Sys.Date()`"
output: html_document
knit: (function(input_file, encoding) {
    out_dir <- '../docs';
    rmarkdown::render(input_file,
      encoding=encoding,
      output_file=file.path(dirname(input_file), out_dir, 'FigS3_scRNAseq_corr.html'))})
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## TL;DR

You may click on the following links to fast forward to the codes directly producing the Figures:

  - [Supplementary Figure 3a](#FigS3a)
  - [Supplementary Figure 3b](#FigS3b)
  - [Supplementary Figure 3c](#FigS3c)
  - [Supplementary Figure 3d](#FigS3d)

Back to [index page](https://ruiyuraywang.github.io/Lin_AAV_Microglia_public/index.html).

## About this Vignette

One of the key questions asked by the reviewers is whether the transduction level (as measured by mScarlet expression mScarlet) is associated with activation of microglia, i.e. relationship between mScarlet level and signature gene. To address this point, we use the Smart-seq2 dataset to make violin plots of signature genes across groups, with overlaying jitter points colored by mScarlet levels.

## Initialization

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratDisk)
  library(tidyverse)
  library(ggplot2)
  library(ggpubr)
  library(patchwork)
  library(RColorBrewer)
})
```

```{r}
querydata <- LoadH5Seurat('../data/AAV_MG_scRNAseq/query_clean.h5Seurat', verbose = F)
query_transferred <- LoadH5Seurat('../data/AAV_MG_scRNAseq/query_transferred.h5Seurat', verbose = F)
```

```{r}
queryaav_sub <- subset(querydata, subset = group == "AAV")
query_transduced <- subset(queryaav_sub, subset = transduction == "Transduced")
```

## Reproducing Plots

### Supplementary Figure 3a {#FigS3a}

```{r}
features_hom = c("P2ry12","P2ry13","Cx3cr1","Csf1r","Lrp1","Ptpre","Lipa","Ets1")
features_hom_lab = c("Sall1", "Sirpa")

vplots <- list()
for(feature in features_hom){
  vplots[[feature]] <- FetchData(queryaav_sub, vars = c("treatment", "transduction", feature, "mScarlet_norm")) %>%
    ggplot(aes_string(x = "treatment", y = feature)) + 
    geom_jitter(aes_string(color = "mScarlet_norm"), size = 1.5, alpha = 1, width = 0.20) + 
    geom_violin(aes_string(fill = "treatment"), alpha = 0.3, width = 0.6) +
    scale_color_distiller('normalized\nmScarlet\nexpression', palette = "Reds", direction = 1) +
    scale_fill_manual(values = brewer.pal(n = 12, name = "Paired")[c(11,7)]) +
    guides(fill = "none") +
    theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.title.y = element_text(size = 15, family = "Arial"),
      axis.text.y = element_text(size = 12, family = "Arial"),
      legend.position = "none",
      legend.title = element_text(size = 10, family = "Arial"),
      legend.text = element_text(size = 9, family = "Arial")
    )
}

for(feature in features_hom_lab){
  vplots[[feature]] <- FetchData(queryaav_sub, vars = c("treatment", "transduction", feature, "mScarlet_norm")) %>%
    ggplot(aes_string(x = "treatment", y = feature)) + 
    geom_jitter(aes_string(color = "mScarlet_norm"), size = 1.5, alpha = 1, width = 0.20) + 
    geom_violin(aes_string(fill = "treatment"), alpha = 0.3, width = 0.6) +
    scale_color_distiller('normalized\nmScarlet\nexpression', palette = "Reds", direction = 1) +
    scale_fill_manual(values = brewer.pal(n = 12, name = "Paired")[c(11,7)]) +
    guides(fill = "none") +
    theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(size = 12, family = "Arial"),
      axis.title.y = element_text(size = 15, family = "Arial"),
      axis.text.y = element_text(size = 12, family = "Arial"),
      legend.position = "none",
      legend.title = element_text(size = 10, family = "Arial"),
      legend.text = element_text(size = 9, family = "Arial"),
    )
}
```
```{r fig.width=6, fig.height=4.5}
ggarrange(vplots[["P2ry12"]], vplots[["P2ry13"]], vplots[["Cx3cr1"]], vplots[["Csf1r"]], vplots[["Sall1"]], vplots[["Sirpa"]], 
          ncol=2, nrow=3, common.legend = TRUE, legend="right")
```

#### Supplementary Figure 3b {#FigS3b}

Alternatively, we could calculate the Spearmann correlation between the expression levels of signature genes and mScarlet transgene. We did not find a correlative relationship between the signature genes and transgene, suggesting that AAV transduction does not induce microglia activation. These are not required by the reviewers but I think they do better at illustrating the points.
```{r}
scplot <- list()
for(feature in c(features_hom, features_hom_lab)){
  scplot[[feature]] <- FetchData(query_transduced, vars = c(feature, "mScarlet_norm")) %>%
    ggscatter(
      x = "mScarlet_norm",
      y = feature,
      xlab = "Normalized mScarlet expression",
      add = c("reg.line"),
      add.params = list(color = "blue", linetype = "dashed"),
      size = .8
    ) +
    stat_cor(
      label.x.npc = 0.70,
      label.y.npc = 0.8,
      label.sep = "\n",
      color = "blue",
      cor.coef.name = "rho",
      method = "spearman",
      size = 2
    ) +
    theme(
      axis.title.x = element_text(size = 9),
      axis.text.x = element_text(size = 12),
      axis.title.y = element_text(size = 15),
      axis.text.y = element_text(size = 12),
    )
}
```
```{r fig.width=5, fig.height=6}
(scplot[["P2ry12"]] | scplot[["P2ry13"]]) / (scplot[["Cx3cr1"]] | scplot[["Csf1r"]]) / (scplot[["Sall1"]] | scplot[["Sirpa"]])
```

#### Supplementary Figure 3c {#FigS3c}

```{r}
features_rea = c("Stat3","Gbp7","Ptgds","Irf7","Irf9","Saa3","Tlr2","Mt1","Mt2","Cd74","Lcn2","Tnf","Il1b","Cst7","Cxcl10",
                 "Ifitm3","Spp1")
features_rea_lab = c("Ccl5","Ccl2")

vplots <- list()
for(feature in features_rea){
  vplots[[feature]] <- FetchData(queryaav_sub, vars = c("treatment", "transduction", feature, "mScarlet_norm")) %>%
    ggplot(aes_string(x = "treatment", y = feature)) + 
    geom_jitter(aes_string(color = "mScarlet_norm"), size = 1.5, alpha = 1, width = 0.20) + 
    geom_violin(aes_string(fill = "treatment"), alpha = 0.3, width = 0.6) +
    scale_color_distiller('normalized\nmScarlet\nexpression', palette = "Reds", direction = 1) +
    scale_fill_manual(values = brewer.pal(n = 12, name = "Paired")[c(11,7)]) +
    guides(fill = "none") +
    theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.title.y = element_text(size = 15, family = "Arial"),
      axis.text.y = element_text(size = 12, family = "Arial"),
      legend.position = "none",
      legend.title = element_text(size = 10, family = "Arial"),
      legend.text = element_text(size = 9, family = "Arial")
    )
}

for(feature in features_rea_lab){
  vplots[[feature]] <- FetchData(queryaav_sub, vars = c("treatment", "transduction", feature, "mScarlet_norm")) %>%
    ggplot(aes_string(x = "treatment", y = feature)) + 
    geom_jitter(aes_string(color = "mScarlet_norm"), size = 1.5, alpha = 1, width = 0.20) + 
    geom_violin(aes_string(fill = "treatment"), alpha = 0.3, width = 0.6) +
    scale_color_distiller('normalized\nmScarlet\nexpression', palette = "Reds", direction = 1) +
    scale_fill_manual(values = brewer.pal(n = 12, name = "Paired")[c(11,7)]) +
    guides(fill = "none") +
    theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(size = 12, family = "Arial"),
      axis.title.y = element_text(size = 15, family = "Arial"),
      axis.text.y = element_text(size = 12, family = "Arial"),
      legend.position = "none",
      legend.title = element_text(size = 10, family = "Arial"),
      legend.text = element_text(size = 9, family = "Arial"),
    )
}
```
```{r fig.width=6, fig.height=6}
ggarrange(vplots[["Tnf"]], vplots[["Il1b"]], vplots[["Cd74"]], vplots[["Cst7"]], vplots[["Spp1"]], vplots[["Irf9"]], vplots[["Ccl2"]], vplots[["Ccl5"]], 
          ncol=2, nrow=4, common.legend = TRUE, legend="right")
```

#### Supplementary Figure 3d {#FigS3d}

```{r}
scplot <- list()
for(feature in c(features_rea, features_rea_lab)){
  scplot[[feature]] <- FetchData(query_transduced, vars = c(feature, "mScarlet_norm")) %>%
    ggscatter(
      x = "mScarlet_norm",
      y = feature,
      xlab = "Normalized mScarlet expression",
      add = c("reg.line"),
      add.params = list(color = "blue", linetype = "dashed"),
      size = .8
    ) +
    stat_cor(
      label.x.npc = 0.70,
      label.y.npc = 0.8,
      label.sep = "\n",
      color = "blue",
      cor.coef.name = "rho",
      method = "spearman",
      size = 2
    ) +
    theme(
      axis.title.x = element_text(size = 9),
      axis.text.x = element_text(size = 12),
      axis.title.y = element_text(size = 15),
      axis.text.y = element_text(size = 12),
    )
}
```
```{r fig.width=5, fig.height=8}
(scplot[["Tnf"]] | scplot[["Il1b"]]) / (scplot[["Cd74"]] | scplot[["Cst7"]]) / (scplot[["Spp1"]] | scplot[["Irf9"]]) / (scplot[["Ccl2"]] | scplot[["Ccl5"]])
```

## Session info

```{r}
sessionInfo()
```